\LARGE \bf ``Those who can't do, teach'': a Teachable Robotic Agent for Children Learning Handwriting

### 2014-09-04 main.tex
WIP

\title{\LARGE \bf
``Those who can't do teach'': Learning Handwriting by Teaching a Robot
}

@@

%\numberofauthors{3} 
%\author{
%\alignauthor
%Deanna Hood\\
%Séverin Lemaignan\\
%Pierre Dillenbourg\\
%    \affaddr{Computer-Human Interaction in\\ Learning and Instruction (CHILI)}\\
%    \affaddr{Ecole Polytechnique Fédérale\\ de Lausanne (EPFL)}\\
%    \affaddr{CH-1015 Lausanne, Switzerland}\\
%    \email{firstname.lastname@epfl.ch}
%\alignauthor
%Rafael Garcia\\
%    \affaddr{Computer Vision and Robotics}\\
%    \affaddr{University of Girona}\\
%    \affaddr{Girona, Spain}
%}
%

@@

Despite the social motivation for the use of a teachable agent for the
engagement and support of children with handwriting difficulties, to date no
such technology is known to have been developed. Presented is a teachable
robotic agent in this context of handwriting skills acquisition. It is
hypothesised that such a system could not just engage an unmotivated student,
but could also present the opportunity for children to experience motor mimicry
during handwriting intervention. It also allows for exploring the potential for
the \emph{learning by teaching} paradigm to be employed in the interaction so as
to stimulate meta-cognition, empathy and increased self-esteem in the child
user.

By leveraging simulated handwriting on a synchronised tablet display, a Nao
humanoid robot with limited fine motor capabilities has been configured as a
suitably embodied handwriting partner. Shape models derived from principal
component analysis of a dataset of letter trajectories are generated and allow
the robot to draw purposefully deformed letters. As a result, the objective of
developing an algorithm learning how to write characters well may be satisfied
by learning the optimal parameters for the model of said characters. Such a
learning algorithm has been developed, capable of incorporating feedback from
users both in terms of which generated letters are the best and from user
demonstrations.

A pilot study, conducted to obtain insight into children's use of the system, is
presented, which has validated the interaction for an in situ experiment
scheduled with a primary school class to evaluate the human-robot interaction
outcomes of the system, the first of its kind.

@@

Handwriting difficulties in children at an early age often negatively affects
the academic performance of the students \cite{Christensen2005} in addition to
their self-esteem being adversely affected \cite{Malloy1995}, causing them to
shy away from expressing what they know \cite{Medwell2008}.
%A conclusion drawn
%in \cite{Hoy2011} is that any handwriting intervention studies considered in the
%systematic review which involved fewer than two practice sessions per week and
%fewer than a total of 20 practice sessions, including homework, were found to
%demonstrate ineffective results. This highlights the necessity to engage
%students in an interaction that will be sustainable over the long-term.

@@

The learning by teaching paradigm, which engages the target student in the act
of teaching another, has been shown to produce motivational, meta-cognition, and
educational benefits in a range of disciplines \cite{Rohrbeck2003}. An
application which is yet to be explored, however, is the use of learning by
teaching in handwriting intervention. One reason why this may be the case is
that in order to engage a child in the learning by teaching paradigm, there must
be an appropriately unskilled peer for them to tutor. While in some cases it may
be appropriate for a peer or teacher to simulate a na\"ive learner, however this
acting is likely to be eventually detected if for an activity where one's skill
level is visually evident - as is the case in handwriting.

Presented is the development of the first known learning agent which is capable
of artificially making mistakes typical of children learning handwriting and
correcting them with external feedback, and therefore exhibiting the potential
to engage children in the learning by teaching paradigm in the context of
handwriting. Furthermore, the learning agent has been embodied in a commercially
affordable humanoid robot by leveraging synchronisation with a tablet display to
simulate fine motor skills otherwise infeasible for the robot.

The significance in the learning by teaching paradigm of observing a teachable
agent respond to what has been taught \cite{Okita2006} is reasoning for using an
embodied agent which is capable of responding to handwriting feedback with a
process rather than simply a product. Therefore investigating the use of a
humanoid capable of executing handwriting trajectories is supported. The
potential for motor mimicry to yield significant improvements in handwriting
interventions \cite{Berninger1997} further supports the case for investigating
the use of a humanoid robot capable of physically demonstrating handwriting
trajectories to its child learning partner.

Section \ref{sec:learningAlgorithm} presents the learning algorithm employed for
developing a learning agent in the context of handwriting, Section
\ref{sec:robotWriting} details the extension of this algorithm to an embodied
robotic learning agent, and Section \ref{sec:experiment} explains the potential
use of such a system in addressing pedagogical research questions, as inferred
from experiments conducted with primary school children.

@@

Because of the objective to use an artificial intelligence algorithm in a
reinforcement learning framework, a parameterisation of letters and their
deformities has been sought such that, depending on the parameters input to the
letter models, different quality shapes can be generated: from bad, to good,
with the help of feedback from the reinforcement learning partner. To this end,
a method which can generate a model from real-world data has been employed in
order to capture said realistic variations.

@@

One approach for a shape model which can appropriately represent realistic
variations in shapes with its parameters is statistical shape modelling, an
application of principle component analysis (PCA), where a linear transform that
de-correlates data vectors is found \cite{Stegmann2002}.

PCA on a set of letter paths captured from a digital pen has been performed,
using the UJI Pen Characters 2 dataset \cite{Llorens2008} with 120 instances of
each letter (2 repetitions from 60 users). While it may be appropriate to
identify the location of the most salient features of the shapes in future work,
the features are currently taken as $n$ uniformly spaced points along the shape
path and arranged into an observation vector presented in (\ref{eq:obsVec}),
where $x_i$ and $y_i$ represent the coordinates of each of the points along the
path.

@@

The projection from the original $2n$ dimensional feature space to the $N$
dimensional feature space is then found as in (\ref{eq:projection}), where ${\bf
p}$ contains the coordinates in the $N$ dimensional space with ${\bf 0}$-origin,
calculated as in (\ref{eq:paramCalc}). The $2n\times N$ matrix ${\bf \Phi }$ is
an orthogonal matrix composed of the eigenvectors ${\bf v}_i$ of the covariance
matrix of the observations which correspond to the largest $N$ eigenvalues
($\lambda_i$), as shown in (\ref{eq:transformCalc}) \cite{Stegmann2002}.

@@

If $N = 2n$, the projection will be nothing more than a change of basis, however
if there is correlation between the points in the observations, there will be
eigenvalues of the covariance matrix which are close to zero, and so removing
the associated eigenvectors from ${\bf \Phi}$ allows for dimensionality
reduction with minimal impact. Indeed, the amount of variance in the dataset
that is explained by each eigenvector ${\bf v}_i$ is $\sqrt{\lambda_i}$, and
therefore the proportion of the variance explained by each dimension/eigenvector
is as shown in (\ref{eq:proportionVar}).

@@

Paths of a particular letter have been clustered using K-means clustering to
group different styles of writing the letter (\emph{allographs}), such that the
parameters in the resulting model are only representative of the specific
allograph of the letter in the dataset. PCA has been performed on all of the
paths of a particular allograph in the dataset at a time, to reduce the $2n$
dimensional space to one of a desired number of dimensions. This process may be
considered to generate a shape model in that (\ref{eq:projection}) may be used
to generate new shapes based on the parameters ${\bf p}$ which are used: ${\bf
p}={\bf 0}$ will yield the mean shape, and variations to each of the $N$ values
in ${\bf p}$ will cause a change in the shape represented by the corresponding
eigenvector (Figure \ref{fig:deviations_sPrint}).

\begin{figure}[thpb]
\centering
\subfigure{
\raisebox{-0.5\height}{\includegraphics[width=0.07\textwidth]{figures/sPrint_mean.png}
}}
\subfigure{
\raisebox{-0.5\height}{\includegraphics[width=0.38\textwidth]{figures/sPrint.png}
}}
\caption[The mean shape and the effect of varying the first five parameters of
the shape model derived from PCA from the dataset of print `s'
shapes.]{\label{fig:deviations_sPrint}The mean shape (left) and the effect
(right) of varying the first five parameters (each row) of the shape model
derived from PCA from the first cluster of `s' shapes, interpreted as the print
allograph. The percentage of the total variance in the dataset explained by each
parameter is shown on the appropriate row. }

@@

Interestingly, the parameters which result from the unsupervised statistical
shape analysis, to some extent, represent variations which could intuitively be
identified by a manual parameterisation: the gap in the top half of the letter
compared to the bottom half, the width of the top half of the letter compared to
the bottom half, the width of the overall shape, etc., and as a result the shape
modelling has produced the desired outcomes in terms of parameterising the
shapes in ways which are able to be used to deform them meaningfully.

@@

After determining the appropriate number of dimensions (N) of the data
representing the majority of the variance observed within different samples of a
letter in the dataset, new letters can be generated by varying the parameter
values in this N dimensional space in accordance with (\ref{eq:projection}). By
choosing parameter values which lie within the observed range in the dataset, it
is possible to produce letters which are more likely to be reasonable looking.
When the parameter values are outside of the range observed in the dataset, they
are less likely to represent shapes from the dataset of reasonable,
adult-written letters, and as a result are more likely to represent poor shapes.
Figure \ref{fig:sampleLetters} illustrates sample letters generated from the
models of `e' and `g' - made from 120 observations with 70 points - by selecting
random values for the first 5 parameters from a distribution with standard
deviation of $3\sqrt\lambda_i$ rather than the $\sqrt\lambda_i$ standard
deviation observed in the dataset.

@@

It is therefore feasible that `poor' letters can be produced using the shape
model generated from a dataset of adults' writing with PCA. In
\cite{Chandra2013}, it was found that children aged 4-6 participating in a
handwriting peer tutoring pilot study most often made mistakes classified as
inappropriate \emph{internal proportions} or \emph{global deformations}, and it
would be reasonable to consider the shapes in Figure \ref{fig:sampleLetters} in
the same general categories. Consequently, while using a database of children's
letters when available may yield potential for parameterising other mistakes
made by children, such as merging strokes or decomposing a shape into multiple
subparts inappropriately, sufficient capabilities for the shape model to
generate shapes given parameter variations which are appropriate for simulating
a na\"ive learning in the context of handwriting have been demonstrated.

@@

In the same way that the shape model of a particular letter can be used to take
parameters and generate a letter, in the case of the statistical shape model
presented in Section \ref{sec:writingGeneration}, it may also be used to take a
letter and determine its parameters given the model. The potential for the
parameters of a user-drawn shape to be obtained from the shape model have been
capitalised on to incorporate the capabilities for the learning algorithm to
adapt to the user's feedback via demonstration.

The statistical shape model may be used to determine the parameters of a
demonstration shape by projecting the features of the observed shape into the
lower-dimensional space determined by the model. Mathematically, given a
demonstration ${\bf x}$, the associated parameters may be determined as in
(\ref{eq:param}).

\begin{equation}\label{eq:param}
{\bf p}_{demo} ={\bf \Phi}^T ({\bf {x}-\bar{x}})
\end{equation}

The parameters which are determined for a shape will not reconstruct the shape
exactly, but rather will reconstruct the closest point in the N dimensional
space which the eigenvectors span. With increasing value of N, the reconstructed
shape will be closer to the original shape, until N is equal to the number of
dimensions of the original shape, at which point there is no error incurred by
`projecting' onto the N dimensional space as the space is the same (but perhaps
with a different basis).

For the statistical shape model employed by the system, the method for
responding to user demonstrations is to move the learning algorithm's parameters
towards those of the demonstration to some extent. In the results presented in
this work, \[{\bf p}^{(k+1)} = {\bf p}^{(k)} + ({\bf p}_{demo}-{\bf
p}^{(k)})\times\alpha\] has been used, where ${\bf p}$ is the parameter vector
and $k$ is the time step, and $\alpha$ is the learning rate, between 0 and 1.
While it is possible that the resulting ${\bf p}^{(k+1)}$ yields an unacceptable
shape from combining ${\bf p}^{(k)}$ and ${\bf p}_{demo}$ which individually
yield acceptable shapes, what is significant about the proposed method for
adapting to user demonstration is that even it does not get stuck at such poor
shapes: with further demonstrations of the same letter, the system would
eventually approach it. It remains to be seen what the impact of such a poor
in-between shape being drawn by the teachable agent would have on the
interaction with the user, however, and how - if necessary - to avoid such an
event mathematically.

@@

@@

In order to develop a teachable agent that is appropriate for engaging a child
through the learning by teaching paradigm - that is, that it is an embodied
agent, capable of eliciting motor mimicry, and capable of causing its teacher to
reflect on the process that has been used to respond to handwriting feedback -
capabilities for the robot to engage in handwriting have been established.

@@

Because of the challenges in getting the Nao to write trajectories smoothly
because of its inherent ability to only reach discrete points in space, it has
been implemented instead that the robot uses simulated handwriting, as an
opportunity to have finer control over the robot's writing. The necessary
components to accomplish such are to:

\begin{enumerate}
    \item develop capabilities for the robot to trace handwriting,

    \item synchronise the robot's movements with the display of the writing on a
        tablet so that it seems that the robot is causing the handwriting to
        display because of its actions, and

    \item integrate the writing behaviour with the handwriting learning
        algorithm presented in Section \ref{sec:learningAlgorithm} to create a
        learning by teaching interaction with the agent. 
\end{enumerate}

These steps are presented in the sections that follow.

@@

As previously mentioned, the robot used in this system does not have
capabilities to reach all points in space due to its physical limitations
(finite degrees of freedom). Using simulated handwriting will not do anything to
fix this physical limitation, but it provides a context where the path traced by
the robot can be displayed smoother than it would have otherwise appeared if
written with a writing instrument. For believability, however, it is still
necessary that the robot's trajectory is close enough to that which is
displayed. As such, the requirements on the robot's fine motor skills are
lessened but not removed.

Aldebaran's NaoQi API has been used for the inverse kinematics of the trajectory
following. It provides the advantage of allowing the whole body to be used in
the inverse kinematics rather than simply the arm, which yields to an increased
working space of the robot. However, as motion planning is performed with
respect to the centre of the hand of the robot as opposed to the tip of a
writing instrument, restrictions on the orientation of the hand must be put in
place such that the writing instrument remains with the appropriate orientation
to the writing surface.

When pretending to write with a writing instrument, the requirement that the
instrument tip match the trajectory is still significant, and therefore any
change in orientation of the hand is exaggerated at the writing tip.

Rather than have the robot write with a writing instrument on a horizontal
surface, by instead having it point at a vertical writing surface to cause the
trajectory to appear (as in Figure \ref{fig:naoWriting}), several advantages are
presented:

@@

    \item Concerns about whether or not the child would start mimicking the
        robot's incorrect (inhuman) writing style are mitigated. However, there
        is a possibility that this would also reduce the potential for students
        to benefit from motor imitation, although it is hypothesised that motor
        imitation is caused more from the overall concept of the trajectory
        being shown than from the pen grip being used.

    \item Perhaps most significantly, the accuracy of the matching of the
        robot's motion with the trajectory displayed on the tablet is not as
        critical, because the finger is not expected to touch the tablet exactly
        at the trajectory point, in contrast to the pen tip. This means that
        orientation degrees of freedom can be freed without significant adverse
        effects.

@@

\begin{figure}[thpb]
     \begin{center}
        %\subfigure{
        %    \includegraphics[width=0.4\textwidth]{figures/naoWriting1.png}
        %}%
        \subfigure{
            \includegraphics[width=0.3\textwidth]{figures/naoWriting2.png}
        }%
%
    \end{center}
    \caption{A demonstration of the robot simulating the drawing of a word
        with its finger, synchronised with the tablet, communicating over ROS.
     }%

@@

Therefore the system presented utilises the context where the robot is
simulating handwriting by pointing at the tablet\footnote{Teachers interviewed
for their feedback on the system advised that children are asked to draw letters
in the air in a similar manner as part of their handwriting education.}.

\subsection{Synchronisation with the Tablet Trajectory Display}

After getting the robot to trace writing trajectories, the next step towards
simulated handwriting is getting the trajectories of the robot's `writing' to
display on an Android tablet instead of being displayed in the RViz
visualisation tool running on the computer.

ROS has been used to connect the devices used in the system, including the
Android tablet\footnote{For more information about ROS on Android devices see
\url{http://wiki.ros.org/android}}. As a result, aspects of the networking
between the tablet and the robot, such as the overheads associated with
connections, ports, message formats, publishers/subscribers, etc. have already
been implemented.

An Android application has been developed to receive the trajectory message over
a ROS topic and add each point of the trajectory one by one as a frame of an
animation. Synchronisation between the tablet and the robot has been achieved by

@@

    \item reducing the number of points along the trajectory passed to the
        robot's motion planner, as the more points, the more the robot gets out
        of synch with the requested timings, and

@@

%\subsection{Integration and Synchronisation}

To instruct the robot where to write, the robot has been configured to detect a
particular fiducial marker, a \emph{chilitag}\footnote{See
\url{https://github.com/chili-epfl/ros_markers} for more information on the
fiducial marker detection node used.}, with the camera located in its head and
use that as the origin of the writing surface (Figure
\ref{fig:tabletDetection}). As publishing the robot's camera has been identified
as a computationally expensive task that can disrupt the robot's timing of
synchronised writing, it is only done when necessary and the tablet is assumed
to be stationary during the writing process.

>>>>

%(DH) is it too soon to claim learning by teaching benefits in the title...? 
%(DH) the paper is not about "learning handwriting", it's about the robot.
\title{\LARGE \bf
``Those who can't do, teach'': a Teachable Robotic Agent for Children Learning Handwriting
}

@@

% \numberofauthors{3} 
% \author{
% \alignauthor
% Deanna Hood\\
% Séverin Lemaignan\\
% Pierre Dillenbourg\\
%    \affaddr{Computer-Human Interaction in Learning and Instruction Laboratory (CHILI)}\\
%    \affaddr{École Polytechnique Fédérale\\ de Lausanne (EPFL)}\\
%    \affaddr{CH-1015 Lausanne, Switzerland}\\
%    \email{firstname.lastname@epfl.ch}
% \alignauthor
% Rafael Garcia\\
%    \affaddr{Computer Vision and Robotics Group (ViCOROB)}\\
%    \affaddr{University of Girona}\\
%    \affaddr{Girona, Spain}
% }

%(DH) my EPFL email address won't work for much longer....

@@

%Despite the social motivation for the use of a teachable agent for the
%engagement and support of children with handwriting difficulties, to date no
%such technology is known to have been developed.

%(DH) abstract probably too long...
%(DH) do we have to be specific somewhere about what stage of handwriting we're targeting..?

Presented is a teachable humanoid agent in the context of handwriting skills
acquisition. It is hypothesised that such a system could not just engage an
unmotivated student, but could also present the opportunity for children to
experience the typical benefits experienced during human-led handwriting
interventions, such as motor mimicry. This system allows for exploring the
potential for the \emph{learning by teaching} paradigm to be employed in the
interaction, so as to stimulate meta-cognition, empathy and increased
self-esteem in the child user.

By leveraging simulated handwriting on a synchronised tablet display, a Nao
humanoid robot with limited fine motor capabilities has been configured as a
suitably embodied handwriting partner. Shape models derived from principal
component analysis of a dataset of letter trajectories are generated and allow
the robot to draw purposefully deformed letters. Learning how to write
characters well is then achieved by incorporating feedback from user
demonstrations, which allows the system to learn the optimal parameters for the
appropriate shape models.

An in situ study has been conducted with primary school classes to obtain
insight into children's use of the novel system. 
%The technical components of the system have been validated, in that children 
Children aged 6-8 were successfully 1) led to believe that the robot was
actually writing on the tablet, and 2) able to improve the robot's writing to a
level which they were satisfied with. Furthermore, the system represents a
significant step towards an innovative use for robotics which addresses a
widespread and socially meaningful challenge in education, and provides context
for robotics in education beyond the well-explored use of teaching coding- and
robotics-related subjects.

@@

Handwriting difficulties in children at an early age often negatively affect
the academic performance of the students \cite{Christensen2005}, in addition to
their self-esteem being adversely affected \cite{Malloy1995}, causing them to
shy away from expressing what they know \cite{Medwell2008}.
%A conclusion drawn
%in \cite{Hoy2011} is that any handwriting intervention studies considered in the
%systematic review which involved fewer than two practice sessions per week and
%fewer than a total of 20 practice sessions, including homework, were found to
%demonstrate ineffective results. This highlights the necessity to engage
%students in an interaction that will be sustainable over the long-term.

@@

The \emph{learning by teaching} paradigm, which engages the target student in
the act of teaching another, has been shown to produce motivational,
meta-cognitive, and educational benefits in a range of disciplines
\cite{Rohrbeck2003}. An application which is yet to be explored, however, is the
application of learning by teaching to handwriting intervention. One reason why
this has not been investigated to date is that in order to engage a child in the
learning by teaching paradigm, there must be an appropriately unskilled peer for
them to tutor. In the context where the target child is the lowest performer in
their class, this may be a constraint. In some cases it may be appropriate for a
peer or teacher to simulate a na\"ive learner for the target child to teach,
however for an activity where one's skill level is visually evident - as is the
case in handwriting - this acting is likely to be eventually detected. As such,
there is motivation for the use of a teachable agent which can be configured for
a variety of skill levels, and for which children do not have preconceptions
about its handwriting ability.

Presented is the development of the first known learning agent which is capable
of artificially making mistakes typical of children learning handwriting, and
correcting them with external feedback. Such a system therefore exhibits the potential
to engage children in the learning by teaching paradigm in the context of
handwriting.

There are three main points which have motivated the use of a robot as the
teachable agent in the presented interaction:

\begin{enumerate}
    \item the engagement and ``protégé effect'' elicited by teaching an embodied
        agent is anticipated to be stronger than that associated with a
        tablet-based application, which children are increasingly recognising as
        disposable technology,

    \item observing a teachable agent respond to what it has been taught
        \cite{Okita2006} is a significant step in the learning by teaching
        paradigm, and therefore it may be significant for the agent to respond
        with a writing \emph{process}, rather than simply a product, and

    \item the potential for motor mimicry to yield significant improvements in
        handwriting interventions \cite{Berninger1997} further supports the case
        for investigating the use of a humanoid robot capable of physically
        demonstrating handwriting trajectories to its child learning partner.
\end{enumerate}

Consequently, the learning agent developed has been embodied in a commercially
affordable Nao humanoid robot. Its writing capabilities have been implemented by
leveraging synchronisation with a tablet display to simulate fine motor skills
otherwise infeasible for the robot.

Section \ref{sec:learningAlgorithm} presents the novel work in the area of
artificial intelligence to develop a learning algorithm suitable for a learning
agent in the context of handwriting.  Section \ref{sec:robotWriting} details the
extension of this algorithm to an embodied robotic learning agent, including the
novel approach for achieving simulated fine motor skills on commercially
affordable humanoid robots.  Section \ref{sec:experiment} explores the novel
contributions made to the study of human-robot interaction, in discussing the
use of the system with primary school children and its potential as a tool for
addressing wider pedagogical research questions in education.

@@

Because of the objective to use an artificial intelligence algorithm in a
reinforcement learning framework, a parameterisation of letters and their
deformities has been sought. Depending on the parameters input to the letter
models, different quality shapes can be generated. The system can therefore be
configured to improve its writing by modifying the parameters used to generate
the letters, based on feedback from the reinforcement learning partner. To this
end, a method which can generate a model from real-world data has been employed
in order to capture realistic variations in letters.

@@

One approach for a shape model which can appropriately represent realistic
variations in shapes with changes to its parameters is statistical shape
modelling. Statistical shape modelling is an application of principle component
analysis (PCA), where a linear transform which de-correlates data vectors is
found \cite{Stegmann2002} and allows for dimensionality reduction.

PCA has been performed on a set of letter paths captured from a digital pen,
using the UJI Pen Characters 2 dataset \cite{Llorens2008} with 120 instances of
each letter (2 repetitions from 60 adult users). While it may be appropriate to
identify the location of the most salient features of the shapes in future work,
the features are currently taken as $n$ uniformly spaced points along the shape
path as salient features are not guaranteed from children's handwriting. The
points are arranged into an observation vector presented in (\ref{eq:obsVec}),
where $x_i$ and $y_i$ represent the coordinates of each of the points along the
path.

@@

The projection from the original $2n$-dimensional feature space $(n=70)$ to an
$N$-dimensional space is found as in (\ref{eq:projection}), where ${\bf p}$
contains the coordinates in the $N$ dimensional space with ${\bf 0}$-origin,
calculated as in (\ref{eq:paramCalc}). The $2n\times N$ matrix ${\bf \Phi }$ is
an orthogonal matrix composed of the eigenvectors ${\bf v}_i$ corresponding to
the largest $N$ eigenvalues ($\lambda_i$) of the covariance matrix of the
observations, as shown in (\ref{eq:transformCalc}) \cite{Stegmann2002}. As such,
each shape may be approximated by the mean shape plus a sum of the top $N$
eigenvectors, weighted by the parameter vector $p$.

@@

If there is correlation between the points in the observations, there will be
eigenvalues of the covariance matrix which are close to zero. As such, removing
the associated eigenvectors from ${\bf \Phi}$ allows for dimensionality
reduction with minimal impact. Indeed, the amount of variance in the dataset
that is explained by each eigenvector ${\bf v}_i$ is $\sqrt{\lambda_i}$, and
therefore the proportion of the variance explained by each dimension/eigenvector
is as shown in (\ref{eq:proportionVar}).

@@

For the dataset presented in Figure \ref{fig:deviations_sPrint}, the
eigenvectors associated with the 5 largest eigenvalues explain 90.0\% of the
variance in the dataset, illustrating the capability of the statistical shape
modelling approach to produce compact parameterisation of shapes.

Clusters within paths of a particular letter have been identified using K-means
clustering to group different styles of writing the letter (\emph{allographs}),
such that the parameters in the resulting model are only representative of the
specific allograph of the letter in the dataset. 
% * <deanna.m.hood@gmail.com> 2014-09-03T13:05:23.576Z:
%
% in the end this was pointless so maybe not worth mentioning
%
PCA has been performed on all of the paths of a particular allograph in the
dataset at a time, to reduce the $2n$ dimensional space to one of a desired
number of dimensions.

(\ref{eq:projection}) may be used
to generate new shapes based on the parameters ${\bf p}$ which are used: ${\bf
p}={\bf 0}$ will yield the mean shape, and variations to each of the $N$ values
in ${\bf p}$ will cause a change in the shape represented by the corresponding
eigenvector (Figure \ref{fig:deviations_sPrint}).

\begin{figure}[thpb]
\centering
\subfigure{
\raisebox{-0.5\height}{\includegraphics[width=0.05\textwidth]{figures/sPrint_mean.png}
}}
\subfigure{
\raisebox{-0.5\height}{\includegraphics[width=0.35\textwidth]{figures/sPrint_top3.png}
}}
\caption[The mean shape and the effect of varying the first three parameters of
the shape model derived from PCA from the dataset of print `s'
shapes.]{\label{fig:deviations_sPrint}The mean shape (left) and the effect
    (right) of varying the first three parameters (each row) of a shape model.
    The presented model has been derived from PCA on the first cluster of `s'
    shapes, interpreted as the print allograph. The extent to which each
parameter is varied is dependent on the eigenvector $\lambda$ corresponding to
the parameter's eigenvector. The percentage of the total variance in the dataset
explained by each parameter is shown to the left of the corresponding row. }

@@

Interestingly, although the parameters are the result of an \emph{unsupervised}
statistical shape analysis, they still represent variations which could have
been intuitively identified by a manual parameterisation. For example, for the
model shown in Figure \ref{fig:deviations_sPrint}, the parameters may represent
the width of the top half of the letter compared to the bottom half, the width
of the overall shape, etc. Consequently, the shape modelling process has
produced the desired outcomes of parameterising the shapes in ways which are
able to be used to deform them meaningfully.

@@

After determining the appropriate number of dimensions ($N$) for the shape
models, new letters can be generated by varying the parameter values in this
$N$-dimensional space in accordance with (\ref{eq:projection}). By choosing
parameter values which lie within the observed range in the dataset, it is
possible to produce letters which are more likely to be reasonable looking.
When the parameter values are outside of the range observed in the dataset, they
are less likely to represent shapes from the dataset of reasonable,
adult-written letters, and as a result are more likely to represent poor shapes.
Figure \ref{fig:sampleLetters} illustrates sample letters generated from the
models of `e' and `g' - made from 120 observations each, with 70 sample points -
by selecting random values for the first 5 parameters from a distribution with
standard deviation of $3\sqrt\lambda_i$ rather than the $\sqrt\lambda_i$
standard deviation observed in the dataset.

@@

In \cite{Chandra2013}, it was found that children aged 4-6 years participating
in a handwriting peer tutoring pilot study most often made mistakes classified
as inappropriate \emph{internal proportions} or \emph{global deformations}. It
would be reasonable to consider the shapes in Figure \ref{fig:sampleLetters} in
the same general categories. Using a database of children's letters when
available may yield potential for parameterising other mistakes made by
children, such as merging strokes or decomposing a shape into multiple subparts
inappropriately. Nevertheless, sufficient capabilities for the shape model to
generate `poor' letters which are appropriate for simulating a na\"ive learning
in the context of handwriting have been demonstrated, using the shape model
generated with PCA from a dataset of only adults' writing.

@@

It was shown in Section \ref{sec:writingGeneration} how the statistical shape
model of a particular letter may convert a set of parameters into a generated
letter of the same allograph. Similarly, the model may also be used to convert a
letter into its parameters, given the model. The parameters of user-drawn
letters may therefore be used in order to incorporate the capabilities for the
learning algorithm to adapt to the user's feedback via demonstration.

The statistical shape model may be used to determine the parameters of a
demonstration shape by projecting the features of the observed shape into the
lower-dimensional space determined by the model. Mathematically, given a
demonstration ${\bf x}_{demo}$, the associated parameters may be determined as in
(\ref{eq:param}).

\begin{equation}\label{eq:param}
{\bf p}_{demo} ={\bf \Phi}^T ({\bf{x}_{demo}-\bar{x}})
\end{equation}

The parameters which are determined for a shape in (\ref{eq:param}) will not
reconstruct the shape exactly if $N<2n$, but rather will reconstruct the closest
point in the $N$-dimensional space which the eigenvectors span. $N=10$ has been
used for the work presented.

For the statistical shape model employed by the system, the method for
responding to user demonstrations is to move the learning algorithm's parameters
towards those of the demonstration to some extent. In the results presented in
this work, \[{\bf p}^{(k+1)} = {\bf p}^{(k)} + ({\bf p}_{demo}-{\bf
p}^{(k)})\times\alpha\] has been used, where ${\bf p}$ is the parameter vector
at time step $k$, and $\alpha$ is the learning rate, between 0 and 1.  It is
possible that parameters ${\bf p}^{(k)}$ and ${\bf p}_{demo}$, which
individually yield acceptable shapes, may produce parameters ${\bf p}^{(k+1)}$
which yield an unacceptable shape. The proposed method for adapting to user
demonstration would not get stuck at such poor shapes, however: with further
demonstrations of the same letter, the system would eventually approach it. As a
result, this event would not limit the interaction proposed in Section
\ref{sec:experiment}, but it may influence the user's perception of the learning
agent. It remains to be seen if it is necessary to avoid such an event
mathematically.

@@

% * <deanna.m.hood@gmail.com> 2014-09-04T09:19:16.057Z:
%
% improve img
%

@@

In order to develop a teachable agent that is appropriate for engaging a child
through the learning by teaching paradigm, 
%- that is, that it is an embodied agent, capable of eliciting motor mimicry,
%and capable of causing its teacher to reflect on the process that has been used
%to respond to handwriting feedback -
capabilities for the robot to engage in handwriting have been established.

@@

%Because of the challenges in getting the Nao to write trajectories smoothly
%because of its inherent ability to only reach discrete points in space, 
A fine level of control over what the robot is writing is necessary in the
proposed application of a learning agent in the context of handwriting. Because
of the limited fine motor skills possible with such an affordable robot, in
addition to the absence of force feedback and other technical necessities, the
Nao has been configured to use simulated handwriting with a synchronised tablet
to achieve this level of control. The necessary components of implementing the
proposed teachable robotic agent are, therefore:

\begin{enumerate}
    \item developing capabilities for the robot to trace handwriting,

    \item synchronising the robot's movements with the display of the writing on a
        tablet so that it seems that the robot is causing the handwriting to
        display because of its actions, and

    \item integrating the writing behaviour with the handwriting learning
        algorithm presented in Section \ref{sec:learningAlgorithm} to create a
        \emph{learning by teaching} interaction with the agent. 
\end{enumerate}

These steps are presented in the sections which follow.

@@

Using simulated handwriting provides an opportunity for the robot's writing to
appear smoother than would be achievable with a writing instrument. However, the
robot's motions must still sufficiently match the displayed trajectory in order
capture the engagement of the child user in the action.

Aldebaran's NaoQi API has been used for the inverse kinematics of the trajectory
following. The Robot Operating System (ROS)\footnote{The ROS stack for Nao is
available at: \url{http://wiki.ros.org/nao_robot}} has been used for integration
of the Nao with external reference frames using the $tf$ transformation library
\cite{Foote2013}, such as the tablet's location.

When using simulated handwriting, it is no longer necessary that the robot
engages in the typical style of handwriting with a writing instrument at a desk.
By instead having the robot point at a vertical writing surface to cause the
trajectory to appear (as in Figure \ref{fig:naoWriting}), several advantages are
presented:

@@

    \item Concerns about whether or not the child would start mimicking the
        robot's incorrect (inhuman) writing style are mitigated. % * <deanna.m.hood@gmail.com> 2014-09-04T09:59:44.707Z:
%
% is this point significant enough?
%

    \item Perhaps most significantly, the accuracy of the matching of the
        robot's motion with the trajectory displayed on the tablet is not as
        critical, because the finger is not expected to touch the tablet exactly
        at the trajectory point, in contrast to the pen tip.
        % This means that orientation degrees of freedom can be freed without significant adverse effects.

@@

\begin{figure}[thpb]
     \begin{center}
        %\subfigure{
        %    \includegraphics[width=0.4\textwidth]{figures/naoWriting1.png}
        %}%
        \subfigure{
            \includegraphics[width=0.3\textwidth]{figures/naoWriting2.png}
        }%
%
    \end{center}
    \caption{A demonstration of the robot simulating the writing of a word
        with its finger. The motion of the robot is synchronised with the display of the tablet, communicating over ROS.
     }%

@@

Therefore, the system presented utilises the context where the robot is
simulating handwriting by pointing at the tablet\footnote{Teachers interviewed
for their feedback on the system advised that children are asked to draw letters
in the air in a similar manner as part of their handwriting education.}. Because
motion planning is performed with respect to the hand of the robot, rather than
its fingertip, the one or two of the orientation degrees of freedom of the hand
are fixed to keep the finger approximately perpendicular to the writing surface,
depending on the desired accuracy. The remaining free orientation(s), coupled
with the whole-body motion control available, allow for a sufficient working
space for writing on the entire tablet.

\subsection{Synchronisation with the Tablet Trajectory Display}\label{sec:tabletSynch}

After getting the robot to trace writing trajectories, the next step towards
simulated handwriting is getting the trajectories of the robot's `writing' to
display on an Android tablet.

ROS has been used for the communication between the devices used in the system,
including the Android tablet\footnote{For more information about ROS on Android
devices see \url{http://wiki.ros.org/android}}. As a result, aspects of the
networking between the tablet and the robot, such as the overheads associated
with connections, ports, message formats, etc. have been simplified.

An Android application has been developed to receive the trajectory message over
a ROS topic and add each point of the trajectory one by one as a frame of an
animation. Synchronisation between the tablet and the robot has been achieved by:

@@

    \item reducing the number of points along the trajectory passed to the
        robot's motion planner, to improve timing accuracy, and

@@

To instruct the robot where to write, the robot has been configured to detect a
particular fiducial marker, a \emph{chilitag}\footnote{See
\url{https://github.com/chili-epfl/ros_markers} for more information on the
fiducial markers used.}, with the camera located in its head and use that as the
origin of the writing surface (Figure \ref{fig:tabletDetection}). When used in
an interaction with a child, this allows the user to move the tablet as required
for the interaction. As publishing the robot's camera has been identified as a
computationally expensive task which can disrupt the robot's timing of
synchronised writing, it is only done when the robot is not writing, and the
tablet is assumed to be stationary during the writing process.
